---
title: "Take-Home Exercise 02"
date: "12 February 2023"
date-modified: "`r Sys.Date()`"
author: "Xinyee How"
format: html
editor: visual
execute: 
  echo: true
  eval: true
  warning: false
---

## Getting Started

### Installing packages

```{r}
pacman::p_load(tidyverse, tmap, sf, spdep, sfdep, readxl, stringr) 
```

### Importing and preparing data

### Geospatial

```{r}
jakarta <- st_read(dsn = "data", 
                 layer="BATAS_DESA_DESEMBER_2019_DUKCAPIL_DKI_JAKARTA")
```

Since we are only concerned about the first 9 rows, we can remove the rest to reduce the number of variables

```{r}
jakarta <- jakarta %>% dplyr::select(c(0:9))
```

**Setting the correct projection**

```{r}
jakarta <- jakarta %>% st_transform(crs=23845)
st_crs(jakarta)
```

**Removing outer islands of Jakarta (by filtering KAB_KOTA column, removing Kabupaten Kepulauan Seribu)**

```{r}
jakarta <- jakarta %>% filter(`KAB_KOTA` %in% c('JAKARTA BARAT',
                                              'JAKARTA PUSAT',
                                              'JAKARTA SELATAN',
                                              'JAKARTA TIMUR',
                                              'JAKARTA UTARA'))
```

**Inspecting layer**

```{r}
jakarta
```

### Aspatial - Vaccination rates of Jakarta from July 2021 to June 2022 (data taken from every last day of the month)

```{r}
setwd("C:/xinyeehow/IS415-GAA/Take-Home_Ex/Take-Home_Ex02/data")
files.list <- list.files(pattern='*.xlsx')
df.list <- setNames(lapply(files.list, read_excel, sheet="Data Kelurahan"), files.list)
df <- bind_rows(df.list, .id = "id")
```

```{r}
df <- df %>%
      mutate_at("id", str_replace, ".xlsx", "")
```

```{r}
df$id <- as.Date(df$id, format="%Y-%m-%d")
```

**Filtering data**, excluding unwanted zones (only want Jakarta Barat, Jakarta Pusat, Jakarta Selatan, Jakarta Timur, and Jakarta Utara in Wilayah Kota column)

```{r}
df <- df %>% filter(`WILAYAH KOTA` %in% c('JAKARTA BARAT',
                                              'JAKARTA PUSAT',
                                              'JAKARTA SELATAN',
                                              'JAKARTA TIMUR',
                                              'JAKARTA UTARA'))
```

**Filtering other unwanted columns** (only keeping administrative boundaries locations and total vaccination information

```{r}
df <- df %>% dplyr::select(c(0:11))
```

**Inspecting vaccine information**

```{r}
df
```

**Joining aspatial and geospatial information by common columns**

```{r}
jkt_df <- left_join(jakarta, df, by = c("DESA" = "KELURAHAN"))
```

**Plotting chloropleth map for examination - first dose**

```{r}
tm_shape(jkt_df) + 
  tm_polygons("JUMLAH\r\nDOSIS 1")
```

We can see some missing values (in grey) in the plot generated above.

**Identifying missing values**

Let's have a glimpse of all the names of the sub-districts first

```{r}
unique(jkt_df$DESA)
```

```{r}
df_unique <- unique(df$KELURAHAN)
```

```{r}
jakarta_unique <- unique(jakarta$DESA)
```

**Identifying problematic names**

```{r}
(df_unique[!(df_unique %in% jakarta_unique)])
```

```{r}
(jakarta_unique[!(jakarta_unique %in% df_unique)])
```

**Fixing problematic names**

```{r}
df$KELURAHAN[df$KELURAHAN == "KERENDANG"] <- "KRENDANG"
df$KELURAHAN[df$KELURAHAN == "RAWA JATI"] <- "RAWAJATI"
df$KELURAHAN[df$KELURAHAN == "KAMPUNG TENGAH"] <- "TENGAH"
df$KELURAHAN[df$KELURAHAN == "BALE KAMBANG"] <- "BALEKAMBANG"
df$KELURAHAN[df$KELURAHAN == "PINANG RANTI"] <- "PINANGRANTI"
df$KELURAHAN[df$KELURAHAN == "JATI PULO"] <- "JATIPULO"
df$KELURAHAN[df$KELURAHAN == "PAL MERIAM"] <- "PALMERIAM"
df$KELURAHAN[df$KELURAHAN == "KRAMAT JATI"] <- "KRAMATJATI"
df$KELURAHAN[df$KELURAHAN == "HALIM PERDANA KUSUMAH"] <- "HALIM PERDANA KUSUMA"
```

**Checking map again for confirmation**

```{r}
jkt_df <- left_join(jakarta, df, by = c("DESA" = "KELURAHAN"))
```

```{r}
tm_shape(jkt_df) + 
  tm_polygons("JUMLAH\r\nDOSIS 1")
```

Now, we can calculate the rates of first, second and third dose rates for Jakarta. We can do so by adding all of the first dose (in all 12 months from July 2021 to June 2022) and dividing by the target vaccinated population of Jakarta to find out how far Jakarta is from the target.

**First dose:**

```{r}
first_dose <- jkt_df %>%
  group_by(DESA, id) %>%
  summarise(`1st Dose` = sum(`JUMLAH\r\nDOSIS 1`) / SASARAN)
```

**Second dose:**

```{r}
second_dose <- jkt_df %>%
  group_by(DESA, id) %>%
  summarise(`2nd Dose` = sum(`JUMLAH\r\nDOSIS 2`) / SASARAN)
```

**Third dose:**

```{r}
third_dose <- jkt_df %>%
  group_by(DESA, id) %>%
  summarise(`3rd Dose` = sum(`JUMLAH\r\nDOSIS 3`) / SASARAN)
```

## Plotting chloropleth maps

**First dose:**

```{r}
tmap_mode("plot")
tm_shape(first_dose) +
    tm_polygons("1st Dose",
            palette = "Blues") +
  tm_facets(by = "id") +
    tm_borders(alpha = 0.5)
```

[Analysis:]{.underline}

As we can visualise from the plot above, the percentage of targeted vaccine rates have been increasing from the beginning of the study (July 2021) to the end of the study (June 2022). The south of Jakarta managed to reach the target the quickest, with the north of Jakarta only reaching the targeted vaccine rates completely only in February 2022, 4 months after south Jakarta reached the targeted vaccine rates.

**Second dose:**

```{r}
tmap_mode("plot")
tm_shape(second_dose) +
    tm_polygons("2nd Dose",
            palette = "Blues") +
  tm_facets(by = "id") +
    tm_borders(alpha = 0.5)
```

[Analysis:]{.underline}

For the second dose, only a few sub-districts managed to hit the targeted vaccine rates even by the end of our study (June 2022). For the rest of the Jakarta sub-districts, most of them only managed to reach 60-80% of the targeted vaccine rates. It seemed like the progression did not change much after the fifth month of study (November 2021).

**Third dose:**

```{r}
tmap_mode("plot")
tm_shape(third_dose) +
    tm_polygons("3rd Dose",
            palette = "Blues") +
  tm_facets(by = "id") +
    tm_borders(alpha = 0.5)
```

[Analysis:]{.underline}

There were no data available for the third dose of vaccine till the ninth month of study (March 2022). For the remaining four months of the study, there seemed to be very little progression to achieve the targeted vaccine rates for all the sub-districts of Jakarta, except for 2 of the sub-districts. For a more detailed analysis of the third dose, a longer study period is required to observe the trend as 4 months is insufficient.

## Local Gi Analysis

**Filtering out 1 month's data (we will use the most recent data - June 2022)**

```{r}
vaccine <- first_dose %>% 
  cbind(second_dose$`2nd Dose`, third_dose$`3rd Dose`)
```

```{r}
june <- vaccine[vaccine$id == "2022-06-30.xlsx",]
```

**Setting seed - for reproducibility**

```{r}
set.seed(9999)
```

**Deriving contiguity weights - Queen's method**

```{r}
wm_q <- june %>%
  mutate(nb = st_contiguity(geometry),
         wt = st_weights(nb,
                         style = "W"),
         .before = 1) 
```

```{r}
wm_q
```

**Computing local Gi statistics**

```{r}
wm_idw <- june %>%
  mutate(nb = st_contiguity(geometry),
         wts = st_inverse_distance(nb, geometry,
                                   scale = 1,
                                   alpha = 1),
         .before = 1)
```

### Computing Hot Spot and Cold Spot Area Analysis

**First Dose**

```{r}
HCSA_1st <- wm_idw %>% 
  mutate(local_Gi = local_gstar_perm(
    `X1st.Dose`, nb, wt, nsim = 99),
         .before = 1) %>%
  unnest(local_Gi)
HCSA_1st
```

**Visualising first dose, setting the significance level (i.e. p-values less than 0.05) hot spot and cold spot areas**

```{r}
HCSA_1st_sig <- HCSA_1st  %>%
  filter(p_sim < 0.05)
tmap_mode("plot")
tm_shape(HCSA_1st) +
  tm_polygons() +
  tm_borders(alpha = 0.5) +
tm_shape(HCSA_1st_sig) +
  tm_fill("gi_star") + 
  tm_borders(alpha = 0.4)
```

This hotspot and coldspot analysis above confirms our conclusion on the first dose of vaccination rates above, that the south of Jakarta is a hotspot and central/north Jakarta are coldspots.

**Second Dose**

```{r}
HCSA_2nd <- wm_idw %>% 
  mutate(local_Gi = local_gstar_perm(
    `second_dose..2nd.Dose.`, nb, wt, nsim = 99),
         .before = 1) %>%
  unnest(local_Gi)
HCSA_2nd
```

**Visualising second dose, setting the significance level (i.e. p-values less than 0.05) hot spot and cold spot areas**

```{r}
HCSA_2nd_sig <- HCSA_2nd  %>%
  filter(p_sim < 0.05)
tmap_mode("plot")
tm_shape(HCSA_2nd) +
  tm_polygons() +
  tm_borders(alpha = 0.5) +
tm_shape(HCSA_2nd_sig) +
  tm_fill("gi_star") + 
  tm_borders(alpha = 0.4)
```

This hotspot and coldspot analysis above further confirms our conclusion on the second dose of vaccination rates above, that the south of Jakarta is a hotspot and central/north Jakarta are coldspots. There seems to be a lot of similarities with the map generated during the first dose analysis.

**Third Dose**

```{r}
HCSA_3rd <- wm_idw %>% 
  mutate(local_Gi = local_gstar_perm(
    `third_dose..3rd.Dose.`, nb, wt, nsim = 99),
         .before = 1) %>%
  unnest(local_Gi)
HCSA_3rd
```

**Visualising third dose, setting the significance level (i.e. p-values less than 0.05) hot spot and cold spot areas**

```{r}
HCSA_3rd_sig <- HCSA_3rd  %>%
  filter(p_sim < 0.05)
tmap_mode("plot")
tm_shape(HCSA_3rd) +
  tm_polygons() +
  tm_borders(alpha = 0.5) +
tm_shape(HCSA_3rd_sig) +
  tm_fill("gi_star") + 
  tm_borders(alpha = 0.4)
```

[Analysis:]{.underline}

For the third dose, the analysis results need to be taken with a pinch of salt, as it was only analysed over a 4-month period. We can conclude that the north of Jakarta still remained as a coldspot for vaccination, and the areas of hotpsot coincides with the higher vaccination rates depicted on the third dose Chloropeth map above.

## Emerging Hot Spot Analysis

Time Series Cube

```{r}
vaccine_st <- vaccine %>% 
  as_spacetime(.loc_col="DESA", .time_col="id")
```

Computing spatial weights using Queen's method

```{r}
vaccine_nb <- vaccine_st %>%
  activate("geometry") %>% 
  mutate(nb = include_self(st_contiguity(geometry)),
         wt = st_inverse_distance(nb, geometry,
                                  scale = 1,
                                  alpha = 1),
         .before = 1) %>% 
  set_wts("wt") %>%
  set_nbs("nb")
```

**Computing Gi values of First Dose (Most taken dose)**

```{r}
gi_stars_EHSA <- vaccine_nb %>%
  group_by(`id`) %>%
  mutate(gi_star = local_gstar_perm(
    `X1st.Dose`, nb, wt, nsim = 99)) %>%
  tidyr::unnest(gi_star)

gi_stars_EHSA
```

## Performing **Mann-Kendall Test**

**Selecting 3 sub-districts to perform test**

Highest first dose rate

```{r}
vaccine %>% 
  top_n(5, `X1st.Dose`) %>%
  arrange(desc(`X1st.Dose`)) %>%
  dplyr::pull(DESA)
```

Lowest first dose rate

```{r}
vaccine %>% 
  top_n(-5,`X1st.Dose`) %>%
  arrange(`X1st.Dose`) %>%
  dplyr::pull(DESA)
```

Middle first dose rate

```{r}
vaccine_ordered <- vaccine %>% arrange(`X1st.Dose`)
vaccine_ordered$DESA[1566]
```

**Computing Mann-Kendall Test**

[Highest first dose rate: HALIM PERDANA KUSUMA]{.underline}

```{r}
cbg_hpk_EHSA <- gi_stars_EHSA %>% 
  ungroup() %>% 
  filter(DESA == "HALIM PERDANA KUSUMA") |> 
  select(DESA, id, gi_star)
```

```{r}
ggplot(data = cbg_hpk_EHSA, 
       aes(x = id, 
           y = gi_star)) +
  geom_line() +
  theme_light()
```

```{r}
cbg_hpk_EHSA %>%
  summarise(mk = list(
    unclass(
      Kendall::MannKendall(gi_star)))) %>% 
  tidyr::unnest_wider(mk)
```

The SL value is the p value, is lower than our 5% significance, indicates that there is a slight upward but insignificant trend for reaching the targeted vaccination rate.

[Middle first dose rate: ROA MALAKA]{.underline}

```{r}
cbg_rm_EHSA <- gi_stars_EHSA %>% 
  ungroup() %>% 
  filter(DESA == "ROA MALAKA") |> 
  select(DESA, id, gi_star)
```

```{r}
ggplot(data = cbg_rm_EHSA, 
       aes(x = id, 
           y = gi_star)) +
  geom_line() +
  theme_light()
```

```{r}
cbg_rm_EHSA %>%
  summarise(mk = list(
    unclass(
      Kendall::MannKendall(gi_star)))) %>% 
  tidyr::unnest_wider(mk)
```

The graph above showed a accelerated decrease in first dose rates being administered on the residents of ROA MALAKA sub-district. This may be due to the fact that most of the residents in MOA MALAKA have gone for their vaccination as soon as they were launched and were very compliant with combating COVID-19.

[Middle first dose rate: BALEKAMBANG]{.underline}

```{r}
cbg_bkb_EHSA <- gi_stars_EHSA %>% 
  ungroup() %>% 
  filter(DESA == "BALEKAMBANG") |> 
  select(DESA, id, gi_star)
```

```{r}
ggplot(data = cbg_bkb_EHSA, 
       aes(x = id, 
           y = gi_star)) +
  geom_line() +
  theme_light()
```

```{r}
cbg_bkb_EHSA %>%
  summarise(mk = list(
    unclass(
      Kendall::MannKendall(gi_star)))) %>% 
  tidyr::unnest_wider(mk)
```

The SL value is the p value, is lower than our 5% significance, indicates that there is a slight upward but insignificant trend for reaching the targeted vaccination rate.

## Emerging Hot Spot Analysis (General - First Dose)

```{r}
EHSA <- emerging_hotspot_analysis(
  x = vaccine_st, 
  .var = "X1st.Dose", 
  k = 1, 
  nsim = 99
)
```

**Visualising emerging hot spot analysis**

```{r}
ggplot(data = EHSA,
       aes(x = classification)) +
  geom_bar()
```

**Visualising geographically**

```{r}
jakarta_EHSA <- jakarta %>%
  left_join(EHSA,
            by = c("DESA" = "location")) %>%
   mutate(`p_value` = replace(`p_value`, `p_value` > 0.05, NA),
          `classification` = ifelse(is.na(`p_value`), NA, `classification`))

tmap_mode("plot")
tm_shape(jakarta_EHSA) +
  tm_fill("classification") +
  tm_borders(alpha = 0.5)
```

**Analysis**

As we can infer from the map plotted out below, the south of Jakarta accounts for majority of the oscilating hotspots for the first dose of vaccination, correlating to our analysis above. The north of Jakarta also correlates to what we saw above, constituting to most of the oscilating coldspots.

## References

With guidance from Prof Kam Tin Seong, and senior Megan Sim Tze Yen
