---
title: "In-Class Exercise 06"
date: "13 February 2023"
date-modified: "`r Sys.Date()`"
author: "Xinyee How"
format: html
editor: visual
execute: 
  echo: true
  eval: true
  warning: false
---

## Installing packages

```{r}
pacman::p_load(tidyverse, tmap, sf, spdep, sfdep) 
```

**Importing geospatial data**

```{r}
hunan <- st_read("data/geospatial/",
                  layer = "Hunan") %>%
  st_transform(crs = 26392)
```

**Importing aspatial data**

```{r}
hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")
```

**Combining Hunan sf dataframe and Hunan_2012 dataframe**

*If want to retain the geometry, have to join the left side of the data frame*

```{r}
hunan_GDPPC <- left_join(hunan,hunan2012)%>% #will auto join by a matching column
  select(1:4, 7, 15) #selecting columns to retain
```

### Plotting chloropleth map

```{r}
tmap_mode("plot") #static map, if want interactive->tmap_mode("view")
tm_shape(hunan_GDPPC)+
  tm_fill("GDPPC", 
          style = "quantile", 
          palette = "Blues",
          title = "GDP Per Capita") +
  tm_layout(main.title = "Distribution of GDP per capita in Hunan",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar() +
  tm_grid(alpha =0.2)
```

## Identify area neighbours

**Calculate contiguity neighbours**

```{r}
cn_queen <- hunan_GDPPC %>%
  mutate(nb= st_contiguity(geometry),
        .before = 1)
```

**Contiguity neighbour list using Rook's method**

```{r}
cn_rook <- hunan_GDPPC %>%
  mutate(nb = st_contiguity(geometry),
         queen = FALSE,
         .before = 1)
```

### Computing contiguity weights

**Queen's method - combining weights**

```{r}
wm_q <- hunan_GDPPC %>%
  mutate(nb = st_contiguity(geometry),
         wt = st_weights(nb),
         .before = 1)
```
