---
title: "Hands On Exercise 01"
date: "`r Sys.Date()`"
format: html
editor: visual
---

For Week 2

## Installing packages

<!--comment like this-->

Install **sf** and **tidyverse**

```{r}
pacman::p_load(sf, tidyverse)
```

## Importing Geospatial Data

**Importing polygon feature data in shapefile format**

```{r}
mpsz = st_read(dsn = "chap01/data/geospatial/master-plan-2014-subzone-boundary-web-shp", 
                  layer = "MP14_SUBZONE_WEB_PL")
```

**Importing polyline feature data in shapefile form**

```{r}
cyclingpath = st_read(dsn = "chap01/data/geospatial/CyclingPath_Jun2022", 
                         layer = "CyclingPathGazette")
```

**Importing GIS data in kml format**

```{r}
preschool = st_read("chap01/data/geospatial/pre-schools-location-kml.kml")
```

## Checking the Content of A Simple Feature Data Frame

**Working with st_geometry()**

```{r}
st_geometry(mpsz)
```

**Working with glimpse()**

```{r}
glimpse(mpsz)
```

**Working with *head()***

```{r}
head(mpsz, n=5) 
```

## Plotting the Geospatial Data

```{r}
plot(mpsz)
```

```{r}
plot(st_geometry(mpsz))
```

```{r}
plot(mpsz["PLN_AREA_N"])
```

## Working with Projection

**Assigning EPSG code to a simple feature data frame**

```{r}
st_crs(mpsz)
```

```{r}
mpsz3414 <- st_set_crs(mpsz, 3414)
```

```{r}
st_crs(mpsz3414)
```

**Transforming the projection of preschool from wgs84 to svy21**

```{r}
preschool3414 <- st_transform(preschool, 
                              crs = 3414)
```

## Importing and Converting Aspatial Data

```{r}
listings <- read_csv("chap01/data/aspatial/listings.csv")
```

```{r}
list(listings) 
```

**Creating a simple feature data frame from an aspatial data frame**

```{r}
listings_sf <- st_as_sf(listings, 
                       coords = c("longitude", "latitude"),
                       crs=4326) %>%
  st_transform(crs = 3414)
```

```{r}
glimpse(listings_sf)
```

## Geoprocessing with sf package

**Buffering**

```{r}
buffer_cycling <- st_buffer(cyclingpath, 
                               dist=5, nQuadSegs = 30)
```

```{r}
buffer_cycling$AREA <- st_area(buffer_cycling)
```
