---
title: "Hands On Exercise 02"
date: "30 January 2023"
date-modified: "`r Sys.Date()`"
author: "Xinyee How"
format: html
editor: visual
execute: 
  echo: true
  eval: true
  warning: false
---

For Week 2

## Installing packages

<!--comment like this-->

Install **sf** and **tidyverse**

```{r}
pacman::p_load(sf, tidyverse)
```

## Importing Geospatial Data

**Importing polygon feature data in shapefile format**

```{r}
mpsz = st_read(dsn = "chap01/data/geospatial/master-plan-2014-subzone-boundary-web-shp", 
                  layer = "MP14_SUBZONE_WEB_PL")
```

**Importing polyline feature data in shapefile form**

```{r}
cyclingpath = st_read(dsn = "chap01/data/geospatial/CyclingPath_Jun2022", 
                         layer = "CyclingPathGazette")
```

**Importing GIS data in kml format**

```{r}
preschool = st_read("chap01/data/geospatial/pre-schools-location-kml.kml")
```

## Checking the Content of A Simple Feature Data Frame

**Working with st_geometry()**

```{r}
st_geometry(mpsz)
```

**Working with glimpse()**

```{r}
glimpse(mpsz)
```

**Working with *head()***

```{r}
head(mpsz, n=5) 
```

## Plotting the Geospatial Data

```{r}
plot(mpsz)
```

```{r}
plot(st_geometry(mpsz))
```

```{r}
plot(mpsz["PLN_AREA_N"])
```

## Working with Projection

**Assigning EPSG code to a simple feature data frame**

```{r}
st_crs(mpsz)
```

```{r}
mpsz3414 <- st_set_crs(mpsz, 3414)
```

```{r}
st_crs(mpsz3414)
```

**Transforming the projection of preschool from wgs84 to svy21**

```{r}
preschool3414 <- st_transform(preschool, 
                              crs = 3414)
```

## Importing and Converting Aspatial Data

```{r}
listings <- read_csv("chap01/data/aspatial/listings.csv")
```

```{r}
list(listings) 
```

**Creating a simple feature data frame from an aspatial data frame**

```{r}
listings_sf <- st_as_sf(listings, 
                       coords = c("longitude", "latitude"),
                       crs=4326) %>%
  st_transform(crs = 3414)
```

```{r}
glimpse(listings_sf)
```

## Geoprocessing with sf package

**Buffering**

```{r}
buffer_cycling <- st_buffer(cyclingpath, 
                               dist=5, nQuadSegs = 30)
```

```{r}
buffer_cycling$AREA <- st_area(buffer_cycling)
```

Total land involved

```{r}
sum(buffer_cycling$AREA)
```

### Point-in-polygon count

**Identify pre-schools located inside each Planning Subzone by using [st_intersects()](https://r-spatial.github.io/sf/reference/geos_binary_pred.html) and [*length()*](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/length) of Base R is used to calculate numbers of pre-schools that fall inside each planning subzone**

```{r}
mpsz3414$`PreSch Count`<- lengths(st_intersects(mpsz3414, preschool3414))
```

**Summary statistics**

```{r}
summary(mpsz3414$`PreSch Count`)
```

**List the planning subzone with the most number of pre-school**

```{r}
top_n(mpsz3414, 1, `PreSch Count`)
```

**Density of pre-school by planning subzone**

```{r}
mpsz3414$Area <- mpsz3414 %>%
  st_area()
```

```{r}
mpsz3414 <- mpsz3414 %>%
  mutate(`PreSch Density` = `PreSch Count`/Area * 1000000)
```

## Exploratory Data Analysis (EDA)

**Distribution of PreSch Density**

```{r}
hist(mpsz3414$`PreSch Density`)
```

```{r}
ggplot(data=mpsz3414, 
       aes(x= as.numeric(`PreSch Density`)))+
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue") +
  labs(title = "Are pre-school even distributed in Singapore?",
       subtitle= "There are many planning sub-zones with a single pre-school, on the other hand, \nthere are two planning sub-zones with at least 20 pre-schools",
      x = "Pre-school density (per km sq)",
      y = "Frequency")
```

Scatter Plot

```{r}
ggplot(data=mpsz3414, 
       aes(y = `PreSch Count`, 
           x= as.numeric(`PreSch Density`)))+
  geom_point(color="black", 
             fill="light blue") +
  xlim(0, 40) +
  ylim(0, 40) +
  labs(title = "",
      x = "Pre-school density (per km sq)",
      y = "Pre-school count")
```
