---
title: "In-Class Exercise 04"
date: "30 January 2023"
date-modified: "`r Sys.Date()`"
author: "Xinyee How"
format: html
editor: visual
execute: 
  echo: true
  eval: true
  warning: false
---

## Installing and Loading the R packages

```{r}
pacman::p_load(maptools, sf, raster, spatstat, tmap)
```

## Importing the spatial data

```{r}
childcare_sf <- st_read("data/child-care-services-geojson.geojson") %>%
  st_transform(crs = 3414)
```

For shapefile --\> have to define the destination (dsn)

SVY21 usually have big numbers (in metres)

Spatstat only can read ppp

so if using sf packages you have to convert spatial dataframe --\> generic dataframe \--\> ppp format (only have coordinate values/no projection headers)

```{r}
sg_sf <- st_read(dsn = "data", layer="CostalOutline")
```

```{r}
mpsz_sf <- st_read(dsn = "data", 
                layer = "MP14_SUBZONE_WEB_PL")
```

```{r}
childcare <- as_Spatial(childcare_sf)
mpsz <- as_Spatial(mpsz_sf)
sg <- as_Spatial(sg_sf)
```

## Converting the Spatial class into generic sp format

```{r}
childcare_sp <- as(childcare, "SpatialPoints")
sg_sp <- as(sg, "SpatialPolygons")
```

## Converting the generic sp format into spatstat's ppp format

```{r}
childcare_ppp <- as(childcare_sp, "ppp")
childcare_ppp
```
