---
title: "Take-Home Exercise 01"
date: "30 January 2023"
date-modified: "`r Sys.Date()`"
author: "Xinyee How"
format: html
editor: visual
execute: 
  echo: true
  eval: true
  warning: false
---

## Exploratory Spatial Data Analysis (ESDA)

### Kernel density maps of functional and non-functional water points using appropriate tmap functions

```{r}
pacman::p_load(maptools, sf, raster, spatstat, tmap)
```

**Kernel density maps on openstreetmap of Osub State, Nigeria**

```{r}

```

**Describe the spatial patterns revealed by the kernel density maps. Highlight the advantage of kernel density map over point map**

---jojo---

## Second-order Spatial Point Patterns Analysis

-   Formulate the null hypothesis and alternative hypothesis and select the confidence level.

    ```{r}

    ```

-   Perform the test by using appropriate Second order spatial point patterns analysis technique.

    ```{r}

    ```

-   With reference to the analysis results, draw statistical conclusions.

    Hello hello

## Spatial Correlation Analysis

-   Formulate the null hypothesis and alternative hypothesis and select the confidence level.

    ```{r}

    ```

-   Perform the test by using appropriate Second order spatial point patterns analysis technique.

    ```{r}

    ```

-   With reference to the analysis results, draw statistical conclusions.

    Hello hello
